Comparing binsegRcpp with other implementations of binary segmentation for changepoint detection

### 2025-05-05 article.Rnw
loss table
blockcpd column

To define the loss functions based on the cumulative sum TODO
\begin{table}
  \begin{tabular}{ccc}
    Parameter & Description & Definition \\
    \hline
    $M=$\code{mean} & model parameter & $T/W$ \\
    \code{var} & model parameter &  $S/W + M(M-2T/W)$

  \end{tabular}
\end{table}

@@

\begin{table}[t!]
\centering
\begin{tabular}{l|ccccc}
  package & binsegRcpp & changepoint & wbs & fpop & ruptures \\
  function & binseg & cpt.mean & sbs & multiBinSeg & Binseg \\
%   > ruptures$version$version
% [1] "1.1.6"
% > sapply(c("binsegRcpp", "changepoint", "wbs", "fpop"), function(p)paste(packageVersion(p)))
%  binsegRcpp changepoint         wbs        fpop 
% "2022.3.29"     "2.2.3"       "1.4" "2019.8.26"
% https://github.com/HaotianXu/changepoints does not have binary segmentation.
  version & 2022.3.29 & 2.2.3 & 1.4 & 2019.8.26 & 1.1.9\\
\hline
  weights & yes & no & no & no & no \\
  max segs & yes & yes & no & yes & yes \\
  dim & one & one & one & multi & multi \\
  correct & yes & no & yes & yes & yes \\
  losses & 5 & 6 & 1 & 1 & 10 \\
  language & C++ & C & C & C++ & Python \\
  % all propose some loss computation.
  storage & STL multiset & arrays & recursion & heap & LRU cache \\
  space & $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$ & $O(S)$ \\
  cumsum & yes & yes & yes & yes & no \\
  best & $O(N\log N)$ & $O(N^3)$ & $O(N\log N)$ & $O(N\log N)$ & $>O(N\log N)$ \\
  worst & $O(N^2)$ & $O(N^3)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$ \\
  CV & yes & no & no & no & no\\
  params & yes & largest & no & no & no \\
\end{tabular}
\caption{\label{tab:overview}
  Properties (rows) of five binary segmentation libraries (columns).
Properties are weights (can specify observation-specific positive weights for the loss), max segs (can specify maximum number of segments parameter), dim (dimensions of input data), correct (does code return correct model parameters?), losses (number of loss functions implemented), language (programming language to implement loss), storage (technique used to store cost of splitting a segment), space (space complexity in number of segments $S$), cumsum (cumulative sum used), best (time complexity in best case data), worst (time complexity in worst case data), CV (support for cross-validation), params (parameters are returned). }
\end{table}

>>>>

To define the loss functions based on the cumulative sum TODO
\begin{table}
  \begin{tabular}{ccc}
    Parameter & Description & Definition \\
    \hline
    $M=$\code{mean} & model parameter & $T/W$ \\
    \code{var} & model parameter &  $S/W + M(M-2T/W)$
  \end{tabular}
  \caption{\label{tab:params}TODO}
\end{table}

\begin{table*}
  \centering
  \begin{tabular}{ccccccc}
    Name & Loss & Distrib. & Center & Scale & Params & Complexity \\
    \hline
    \texttt{mean\_norm} & Square & Normal & yes & no & 1 & $O(S)$ \\
    \texttt{meanvar\_norm} & Neg. Log. Lik. & Normal & yes & yes & 2 & $O(S)$ \\
    \texttt{poisson} & Neg. Log. Lik. & Poisson & yes & yes & 1 & $O(S)$ \\
    \texttt{l1} & Absolute & Laplace & yes & no & 1 & $O(S\log S)$ \\
    \texttt{laplace} & Absolute & Laplace & yes & yes & 2 & $O(S\log S)$ \\
  \end{tabular}
  \caption{\label{tab:losses}
    TODO
  }
\end{table*}

@@

\begin{table}[t!]
  \centering
  \scriptsize
\begin{tabular}{l|cccccc}
  package & binsegRcpp & changepoint & wbs & fpop & ruptures & blockcpd\\
  function & binseg & cpt.mean & sbs & multiBinSeg & Binseg & fit\_blockcpd\\
%   > ruptures$version$version
% [1] "1.1.6"
% > sapply(c("binsegRcpp", "changepoint", "wbs", "fpop"), function(p)paste(packageVersion(p)))
%  binsegRcpp changepoint         wbs        fpop 
% "2022.3.29"     "2.2.3"       "1.4" "2019.8.26"
% https://github.com/HaotianXu/changepoints does not have binary segmentation.
  version & 2022.3.29 & 2.2.3 & 1.4 & 2019.8.26 & 1.1.9 & 1.0.0\\
\hline
  weights & yes & no & no & no & no & no\\
  max segs & yes & yes & no & yes & yes & yes\\
  dim & one & one & one & multi & multi & multi \\
  correct & yes & no & yes & yes & yes \\
  losses & 5 & 6 & 1 & 1 & 10 & 5\\
  language & C++ & C & C & C++ & Python & C++\\
  % all propose some loss computation.
  storage & STL multiset & arrays & recursion & heap & LRU cache & heap \\
  space & $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$ & $O(S)$ & $O(S)$ \\
  cumsum & yes & yes & yes & yes & no & yes\\
  best & $O(N\log N)$ & $O(N^3)$ & $O(N\log N)$ & $O(N\log N)$ & $>O(N\log N)$ & $O(N\log N)$\\
  worst & $O(N^2)$ & $O(N^3)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$\\
  CV & yes & no & no & no & no & no\\
  params & all & one & no & no & no &one\\
\end{tabular}
\caption{\label{tab:overview}
  Properties (rows) of five binary segmentation libraries (columns).
Properties are weights (can specify observation-specific positive weights for the loss), max segs (can specify maximum number of segments parameter), dim (dimensions of input data), correct (does code return correct model parameters?), losses (number of loss functions implemented), language (programming language to implement loss), storage (technique used to store cost of splitting a segment), space (space complexity in number of segments $S$), cumsum (cumulative sum used), best (time complexity in best case data), worst (time complexity in worst case data), CV (support for cross-validation), params (parameters are returned). }
\end{table}
